---
title: "Modeling and prediction for movies"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  word_document: default
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(GGally)
library(knitr)
```

### Load data


```{r load-data}
load("eaca_movies.Rdata")
d1<-as.data.frame(movies)
```



* * *

## Part 1: Data

The data set is a randomly selected group of 651 movies produced and released before 2016.  Since this data is observational and not related to an expirement with random assignment, we can not look for causal relationships.  However, since the movies were randomly selected, we can look for associations and apply them to the population of interest, movies produced and released for consumption in the US market.


* * *

## Part 2: Research question

I'm interested in using the associations in the data set to try to explain the variance in the values of the critics score from the Rotten Tomatoes website.  I will try to develop the best possible multi-variant regression model to accomplish this.  I'll then use the model to predict the critics score for movies produced after 2016 that the model has not seen before.



* * *

## Part 3: Exploratory data analysis

I'll first eliminate variables that don't contribute to my analysis.  The movie's title, director, and featured actors will not be helpful as they would be categorical variables with almost as many levels as observations.  The individual urls for both IMDb and Rotten Tomatoes will also not contain useful information.  The data also contains both ratings and scores for the movies from both websites.  The ratings are just groups of scores, so  I am going to use the scores and drop the ratings to keep all movie ratings as continuous numerical variables.

This code will select the varialbles I want to use, check for and drop any observations with missing data and check the dimensions of the updated data set:

```{r}
# variables to keep 

keep<-c("title_type", "genre","runtime","mpaa_rating", "thtr_rel_year",
        "thtr_rel_month", "thtr_rel_day", "dvd_rel_year", "dvd_rel_month",
        "dvd_rel_day", "imdb_rating", "imdb_num_votes", "audience_score",
        "critics_score", "best_pic_nom",
        "best_pic_win", "best_actor_win", "best_actress_win", "best_dir_win",
        "top200_box")

kept<-d1 %>%
  select(keep)

# remove records with nas

sum(is.na(kept))
cc<-complete.cases(kept)
kept<-kept[cc,]
sum(is.na(kept))
dim(kept)

```

Next, we'll look at summary statistics for the numerical values in the data set:

```{r}

# summary statistics

keptnum<-select_if(kept, is.numeric)

s1<-summary(keptnum$runtime)
s2<-summary(keptnum$thtr_rel_year)
s3<-summary(keptnum$thtr_rel_month)
s4<-summary(keptnum$thtr_rel_day)
s5<-summary(keptnum$dvd_rel_year)
s6<-summary(keptnum$dvd_rel_month)
s7<-summary(keptnum$dvd_rel_day)
s8<-summary(keptnum$imdb_rating)
s9<-summary(keptnum$imdb_num_votes)
s10<-summary(keptnum$audience_score)
s11<-summary(keptnum$critics_score)

Stats<-names(keptnum)

sumtable<-as.data.frame(rbind(s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11))

sumtable<-as.data.frame(cbind(Stats, sumtable))

sumtable<-sumtable %>%
  mutate(Skew=ifelse(Mean>Median, "Right", "Left"))

kable(sumtable)


```


We see from the summary statistics that the IMDb number of votes is exrremely right skewed.  Most of the films are between and hour and a half and two hours, with a few much longer.  The IMDb ratings and Rotten Tomatoes scores are left skewed with the ratings/scores bunching up a bit at the higher end of the range.


The following code will select the numerical values in the data set and plot them in a correlation matrix.  This will allow us to see how these variables correlate with our dependant variable, and help us look for colinearity between independent variables. It also provides a density curve which will help us look at the distributions of the numeric variables.

```{r, fig.width=11, fig.height=13}

# check correlarions for colinearity

keptnum<-select_if(kept, is.numeric)

p1<-ggpairs(keptnum)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))

p1
```


From the plot we see that the IMDb score and the audience score from Rotten Tomatoes have the highest correlation coefficient with the critic's score.  They are also highly correlated with each other, so I will need to use one or the other in my model.  IMDb number of votes is colinear with IMDb rating, so I will use only one of these two as well.



We'll now look at how the critic's score breaks out by some of the categorical variables.

This plot looks at the critic's score by genre:

```{r}
p2<-ggplot(kept, aes(x=factor(genre), y=critics_score, fill=factor(genre)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Critic's Score by Genre")

p2
```

We see from this plot that documentaries and musicals have the highest average scores.

This plot looks at critic's score by the film's rating:

```{r}

p3<-ggplot(kept, aes(x=factor(mpaa_rating), y=critics_score, fill=factor(mpaa_rating)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Critic's Score by Rating")

p3

```

We see from this plot that unrated films have the highest scores on average, and the PG=13 films have the lowest average scores.

Now that we've taken a close look at the data set, it's time to start building our model.


* * *

## Part 4: Modeling

First, I need to decide if I'm going to use the audience score or the IMDb rating in my model.  Based on my EDA, these two independent variables are tightly correlated with each other, so I need to use one or the other in my model.  To select which one, I will regress each against the critic's score and select the on with the highest adjusted R squared.

```{r}

tm1<-lm(critics_score~imdb_rating, data=kept)

summary(tm1)

tm2<-lm(critics_score~audience_score, data = kept)

summary(tm2)

kept$audience_score<-NULL

```

We see that the IMDb rating gives us a better adjusted R squared, so we will proceed using that variable.

I will also exclude the IMDb number of votes variable as it does not have anywhere close to a linear relationship with the dependant variable.

```{r}

qqplot(keptnum$imdb_num_votes, keptnum$critics_score)

kept$imdb_num_votes<-NULL


```


I will now build my model using forward selection with adjusted R squared as the selection metric.  I went with forward selection because it appears from the correlation matrix that I will have one or two variables explaining most of the variance.  I will use the adjusted R squared method as I am more interested in predicting the dependant variable than understanding the significance of the various predictors.

This code will run a linear model for each of the independent variables 
against the critic's score and a assemble a table with the independent variables ranked by adjusted R squared:

```{r}

# create df of independent variables

indvar<-kept[,-12]

# store variable names in vector

varlist <- names(indvar)

# function to run lms

models <- lapply(varlist, function(x) {
  lm(substitute(critics_score ~ i, list(i = as.name(x))), data = kept)
})

# grab adjusted r squared and output ranked in descending order

adjr2<-NULL
options(scipen=999)

for (i in 1:length(models)){
  adjr2[[i]]<-summary(models[[i]])$adj.r.squared
}

results1<-as.data.frame(cbind(varlist, adjr2))

results1<-results1 %>%
  arrange(desc(adjr2))
  
  kable(results1)
  
```

We see that the IMDb rating is the variable that results in the highest adjusted R squared.  This is not a surprise based on the correlation matrix from the EDA.

I continued this process eleven more times, adding an additional variable to the model each time until the adjusted R squared went down instead of up.  The final selected model is as follows:

```{r, fig.width=11, fig.height=13}

final<-lm(critics_score ~ imdb_rating+mpaa_rating+title_type+
                genre+thtr_rel_year+dvd_rel_year+dvd_rel_day+
                best_dir_win+best_pic_nom+runtime+top200_box, data = kept)

summary(final)

```

### Interpretation of Coefficients

From the model summary we see that the intercept is -445, which is the critic's score we would expect if all of the independent variables were set to 0.  The significant numeric variables are the IMDb rating and the year of the theatrical release.  For each 1 point increase in the IMDb rating, we would expect the critic's score to increase 18.2 points on average.  For each increase in year, we would expect the critic's score to decrease .33 points.  This association is interesting as it indicates that newer movies tend to be rated lower than older movies by the contributing critics on the Rotten Tomatoes site.  Of the categorical variables, the genre of the movie returns a significant association if the movie is a drama.  A drama will add 8.54 points to the critic's score relative to the reference level - Action / Adventure.

### Model Diagnostics

I'll now check my model against the conditions for multiple linear regression.  First we'll make sure our residuals have a nearly normal distribution with a mean of 0.

```{r, fig.width=11, fig.height=3}

mean(final$residuals)

par(mfrow=c(1,2))

qqnorm(final$residuals)
hist(final$residuals)

```

Next, we'll check to make sure the numeric independent variables have a linear relationship with our dependant variable.  We've already looked at this in the correlation matrix, but we'll look again here:

```{r}

par(mfrow=c(2,3))

plot(kept$critics_score~kept$runtime)
abline(lm(kept$critics_score~kept$runtime), col="red", lwd=3)

plot(kept$critics_score~kept$thtr_rel_year)
abline(lm(kept$critics_score~kept$thtr_rel_year), col="red", lwd=3)

plot(kept$critics_score~kept$dvd_rel_year)
abline(lm(kept$critics_score~kept$dvd_rel_year), col="red", lwd=3)

plot(kept$critics_score~kept$dvd_rel_day)
abline(lm(kept$critics_score~kept$dvd_rel_day), col="red", lwd=3)

plot(kept$critics_score~kept$imdb_rating)
abline(lm(kept$critics_score~kept$imdb_rating), col="red", lwd=3)

```

Aside from the IMDb rating, these relationships are pretty weak, but they still add a bit of additional information on our dependant variable based on the adjusted R squared method.

Next we'll confirm the constant variability of the residuals by looking at a plot of the residuals vs. fitted values

```{r}

plot(final$residuals~final$fitted.values)

```

The residuals appear randomly distributed around 0, so this condition seems to be met.

Finally, we'll check the independence of the residuals by plotting them against the order of data collection (index).

```{r}

plot(final$residuals)

```

There do not seem to be anything but a random pattern here, so it seems we've met all the conditions for multiple linear regression.  Let's now test the predictive power of the model with some new data.
  
* * *

## Part 5: Prediction

I'll now use the predict function to predict the Rotten Tomatoes critic's score for the following four films.  I collected the data for these movies from the IMDb website (https://www.imdb.com/), the Rotten Tomatoes website (https://www.rottentomatoes.com/), and the Box Office Mojo website (http://www.boxofficemojo.com/).

- The Revenant
- La La Land
- Get Out
- The Space Between us

The first three films all have high critic's scores, however the score for the Revenant seems a bit low for a film that won two Oscars.  The Space Between Us has a very low critic's score - especially compared to the audience score.  I deliberately selected some extreme values to test the predictive powers of the model.  I expect I may lose a few points on the accuracy as a result, but let's see what happens.....

The following code will predict the critics score for the films mentioned above and return a table with the predicted value, upper and lower bounds for the uncertainty based on a 95% confidence interval, and the actual critic's score for the film:

```{r}

# Predict

# The Revenant
# La La Land
# Get Out
# The Space Between Us

newdata<-read.csv("newdata.csv")

preds<-as.data.frame(predict(final, newdata=newdata, interval="confidence", level=.95))

Title<-c("The Revenant", "La La Land", "Get Out", "The Space Between Us")

preds<-as.data.frame(cbind(Title, preds))

names(preds)<-c("Title", "Predicted Score", "Lower 95%", "Upper 95%")

preds$ActualScore<-newdata$critics_score

kable(preds)

```

The model's confidence intervals includes the actual critic's score for two of the four films.  The model only missed badly on "The Space Between Us".  This isn't surprising as the audience metrics from both IMDb and Rotten Tomatoes are not nearly as bad as the critic's score.  Since the IMDb rating was the most significant predictor in the model, the model then over-estimated the critic's score for that film.


* * *

## Part 6: Conclusion

This analysis has shown me that there is a strong connection between film critics opinions and the public at large.  From the data set provided, the variable representing the public's opionion (IMDb score) accounted for the vast majority of the variance in the dependant variable (Rotten Tomatoe's critits score) explained by my model.  There can, however, be significant divergence of opinion, as evidenced by the prediction for "The Space Between Us".  I think it would be interesting to investigate additional data sources for critical reviews to include in subsequent models. 

All of the r code for this analysis can be found here:

https://github.com/lecroc/Linear-Regression-and-Modeling

